Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation

### 2024-01-22 presentation-before-Jan-2024.qmd
change presi

---
title: "Praise the people or praise the place? Upper tail human capital in electrifying Sweden"
subtitle: "Midway Seminar - 24-01-2024"
author: 
  - "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview_wide.png"
  data-background-position: "bottom 10"
  data-background-size: "90%"
  data-background-opacity: "0.4"
bibliography: bibliography.bib
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com/content/research.html)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    include-in-header: 
        - text: <link rel = "shortcut icon" href = "assets/favicon.ico" />

@@

# Research question{#sec-researchquestion background-color="#1C9E77" visibility="uncounted"}

How can we use new kinds of source material and new methods to find out where the skilled workers who built the Swedish electricity sector came from, and where they were trained?

Answering involves:

- 2 data structuring tasks
- 2 classification tasks

@@

## Related literature

**3 main strands:**

::: {.columns}
::: {.column}
*Recent papers making use of individual level biographic data*

- @ford2023not
- @Nekoei2020Herstory

:::
::: {.column}
*The importance of human capital in economic development*

- @Goldin1994
- @Moretti2012
- @Mokyr2017

:::
::: {.column}
*Classification and prediction as part of EH*

- @DahlVedel2024
- @merouani2023innovation
- @Bandiera2020ceo

:::
:::

# Who were the `innovators` or adopters?{#sec-who background-color="#1C9E77" visibility="uncounted"}

@@

@@

[Lund, Karl Gustaf]{style="background-color: #b0c4de"}, [chief engineer]{style="background-color: #d2b48c"}, [born on July 22, 1893]{style="background-color: #ff8c69"}, [in Hille, Gävleborgs län]{style="background-color: #90ee90"}, Sweden, son of [clerk Ferdinand L. and Maria Andersson]{style="background-color: #9370db"}. Married in 1936 to [Sigrid Johansson]{style="background-color: #ffadc1"}. Children: [Ingvar (born 1938)]{style="background-color: #ffdead"}, [Lennart (born 1942)]{style="background-color: #ffdead"}. — Graduated from [Bergsskolan]{style="background-color: #afeeee"} in Filipstad in 1917, specialized studies at the [Royal Institute of Technology (KTH)]{style="background-color: #afeeee"} from 1920 to 1922, studied at the [Institute of Metallurgy]{style="background-color: #afeeee"} and [Stockholm University]{style="background-color: #afeeee"} in 1921-1922. Chemist at [Strömsnäs Järnverks A-B]{style="background-color: #98fb98"} in Degerfors from 1918 to 1920, metallurgist and chemist at [Westinghouse Electric & Manufacturing Co.]{style="background-color: #98fb98"} in East Pittsburgh, PA, USA, from 1923 to 1926 and 1928 to 1929, chief metallurgist at [Laclede Steel Co.]{style="background-color: #98fb98"} in Alton, Illinois, USA, in 1927, furnace and steelworks engineer at [A-B Iggesunds Bruk]{style="background-color: #98fb98"} from 1929 to 1931, site manager at [Gunnebo Bruks Nya A-B, Varbergsverket]{style="background-color: #98fb98"}, since 1931. Member of the municipal executive committee, deputy chairman of the economic department, deputy member of the board of the power plant, chairman of [Varbergs Sparbank]{style="background-color: #ffdab9"}, employer representative in the district council of the county labor board, member of the board of [Varbergs Luftskyddsförening (Varberg Air Protection Association)]{style="background-color: #ffdab9"}, secretary of [Varbergs Högerförening (Varberg Conservative Association)]{style="background-color: #ffdab9"}, chairman of the railway sick fund, and [Plant Society for Small Bird Friends]{style="background-color: #ffdab9"}.

# Where were the `firms` in which these individuals worked?{#sec-where background-color="#1C9E77" visibility="uncounted"}

## Swedish industrial firm catalogue of 1947

::: {.columns}
::: {.column}
![](assets/svensk-industrikalender_1.png){fig-align="center" width=100%}
:::
::: {.column}
The Svensk Industrikalender, published by Sveriges Industriförbund (Sweden’s Industrial Association), is a comprehensive directory of Swedish industrial firms. This calendar was issued annually from 1918 to 2000 and contains information related to Swedish industry. The 1947 edition available on the Project Runeberg website was digitized in April 2012, sourced from the Centrum för Näringslivshis- toria.
:::
:::

## Firms across the country

```{=html}
<div style="display: flex; justify-content: center; align-items: center;">
    <iframe src="https://raw.githack.com/j-jayes/svensk-industrikalender/main/data/maps/all_companies_map.html" width="1400" height="700"></iframe>
</div>
```

# How can we `structure` the data?{#sec-structure background-color="#1C9E77" visibility="uncounted"}

@@

| Step | Process                                   |        | Description                                                           |
|------|-------------------------------------------|-------------|
| 1    | Scrape book data from website             | 🌐➡️📚       | 
| 2    | Split records on each page of a book  | 📚➡️📄       | 
| **3**    | **Structure biographies using schema**    | 📄🔄📄    | 
| **4**    | **Classify record into group**                              | 🔗🔍🌍🎓📚   |
| 5    | Store data for analysis                                | 🔗📚📂       |

@@

## Structure using OpenAI API 🤖

@@

# How can we `classify occupations` according to a schema?{#sec-classifying-occupations background-color="#1C9E77" visibility="uncounted"}

@@

## References

::: {#refs}
:::

>>>>

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
subtitle: "Labour Mobility Lab in Lund"
date: today
author: "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.2"
  data-background-position: "right"
format:
  clean-revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
favicon: html/favicon.ico

bibliography: bibliography.bib

@@

@@

1. **Background: Return Migrants and Occupational Mobility in Sweden** 🌍➡️🇸🇪
    - Return migrants brought back financial capital, often invested locally, but didn't show much occupational upgrading on average.

2. **How are the individuals at the top of the skill distribution different from those who returned to farm?** 
    - Where did they move back to, how did their education change?

# Research question{background-color="#1C9E77"}

How important was **migration and skill adoption** for the **electricity adjascent technology development** in Sweden?

[[Skip to Classifying Occupations]{.button}](#sec-classifying-occupations)

## What I want you to take away from today as researchers

::: {.column width="40%"}
1. NLP tasks that used to be very challenging / time consuming are no longer!

2. We can use the OpenAI API to structure data from free text at very low cost.

3. It is a **General Purpose Technology** that can both process data and also inform our workflow as researchers.

::: {.column width="10%"}
:::

::: {.column width="50%"}
![](assets/world_champtions.jpeg){fig-align="center" width=60%}
:::

# Structuring biographies{background-color="#1C9E77"}

@@

## Vem är Vem? Example biography

![](assets/Karl_Lund.png)

@@

[Lund, Karl Gustaf]{style="background-color: #b0c4de"}, [chief engineer]{style="background-color: #d2b48c"}, [born on July 22, 1893]{style="background-color: #ff8c69"}, [in Hille, Gävleborgs län]{style="background-color: #90ee90"}, Sweden, son of [factory worker Ferdinand L. and Maria Andersson]{style="background-color: #9370db"}. Married in 1936 to [Sigrid Johansson]{style="background-color: #ffadc1"}. Children: [Ingvar (born 1938)]{style="background-color: #ffdead"}, [Lennart (born 1942)]{style="background-color: #ffdead"}. — Graduated from [Bergsskolan]{style="background-color: #afeeee"} in Filipstad in 1917, specialized studies at the [Royal Institute of Technology (KTH)]{style="background-color: #afeeee"} from 1920 to 1922, studied at the [Institute of Metallurgy]{style="background-color: #afeeee"} and [Stockholm University]{style="background-color: #afeeee"} in 1921-1922. Chemist at [Strömsnäs Järnverks A-B]{style="background-color: #98fb98"} in Degerfors from 1918 to 1920, metallurgist and chemist at [Westinghouse Electric & Manufacturing Co.]{style="background-color: #98fb98"} in East Pittsburgh, PA, USA, from 1923 to 1926 and 1928 to 1929, chief metallurgist at [Laclede Steel Co.]{style="background-color: #98fb98"} in Alton, Illinois, USA, in 1927, furnace and steelworks engineer at [A-B Iggesunds Bruk]{style="background-color: #98fb98"} from 1929 to 1931, site manager at [Gunnebo Bruks Nya A-B, Varbergsverket]{style="background-color: #98fb98"}, since 1931. Member of the municipal executive committee, deputy chairman of the economic department, deputy member of the board of the power plant, chairman of [Varbergs Sparbank]{style="background-color: #ffdab9"}, employer representative in the district council of the county labor board, member of the board of [Varbergs Luftskyddsförening (Varberg Air Protection Association)]{style="background-color: #ffdab9"}, secretary of [Varbergs Högerförening (Varberg Conservative Association)]{style="background-color: #ffdab9"}, chairman of the railway sick fund, and [Plant Society for Small Bird Friends]{style="background-color: #ffdab9"}.

Traveled to [Germany]{style="background-color: #e6e6fa"} in 1921, 1923, 1930, and 1936, [Denmark]{style="background-color: #e6e6fa"}, [Czechoslovakia]{style="background-color: #e6e6fa"} in 1921, 1922, and 1923, [Austria]{style="background-color: #e6e6fa"} in 1921, and the [USA]{style="background-color: #e6e6fa"} from 1923 to 1929. Writings: "Some fundamental factors for obtaining sharp thermal curves" (Transactions of the American Society for Steel Treating, co-authored with C. Benedicks and W. H. Dearden, 1925), "Contemporary production of saw blades, circular saws, and machine knives" (Timber Industry, 1931). Hobbies: [hunting]{style="background-color: #ffd700"} and [fishing]{style="background-color: #ffd700"}.

##

![](assets/tidy.jpg)

@@

| Step | Process                                   |        | Description                                                           |
|------|-------------------------------------------|-------------|
| 1    | Scrape book data from website             | 🌐➡️📚       | 
| 2    | Split biographies on each page of a book  | 📚➡️📄       | 
| **3**    | **Translate biographies**                     | 📄🔄📄       | 
| **4**    | **Structure biographies using Schema.org**    | 📄➡️🔗       | 
| **5**    | **Augment data**                              | 🔗🔍🌍🎓📚   |
| 6    | Store data                                | 🔗📚📂       |

@@

## 1. Translate 🤖

@@

# Classifying occupations{#sec-classifying-occupations background-color="#1C9E77"}

@@
